---
title: "BSA"
author: "Tom Saunders"
format: gfm
knitr:
  opts_chunk:
    fig.path: figs/
---

```{r}
#| message: false
library(readxl)
library(lubridate)
library(readr)
library(dplyr)
library(stringr)
library(tidyr)
```

Read in two datasets: the first is for complaints made under the previous BSA Codebook (1 April 2016 to 30 June 2022); the second is for complaints made under the current Codebook (1 July 2022 to present).

```{r}
bsa_prev <- read_excel("data/20230810 AllClosedFormalComplaints.xlsx")
bsa_curr <- read_excel("data/20230807 AllClosedFormalComplaints.xlsx")
```

Convert `Broadcast Date/Time` to date in each dataset. I can't find a way to preserve times.

```{r}
bsa_prev$`Broadcast Date/Time` <- as.numeric(bsa_prev$`Broadcast Date/Time`)
bsa_prev$`Broadcast Date/Time` <- as.Date(bsa_prev$`Broadcast Date/Time`, origin = "1899-12-30")

bsa_curr$`Broadcast Date/Time` <- as.Date(bsa_curr$`Broadcast Date/Time`, origin = "1899-12-30")
```

Join the two datasets.

```{r}
#| message: false
bsa_full <- full_join(bsa_prev, bsa_curr)
```

Rename columns.

```{r}
colnames(bsa_full) <- c("broadcaster", "complaint", "programme", "broadcast_date", "genre",
                        "code", "determination", "decision_date", "majority", "split", 
                        "elect_standards", "elect_notupheld", "elect_upheld", 
                        "fta_standards", "fta_notupheld", "fta_upheld", "paytv_standards",
                        "paytv_notupheld", "paytv_upheld", "radio_standards", "radio_notupheld",
                        "radio_upheld", "standards", "not_upheld", "upheld", "tv_radio")
```

Consolidate individual standards based on previous codebook into 'standards' column introduced for new codebook. Repeat for 'upheld' and 'not upheld' standards. Drop these columns once they are consolidated.

```{r}

bsa_full$standards <- coalesce(bsa_full$elect_standards, bsa_full$fta_standards, bsa_full$paytv_standards, bsa_full$radio_standards, bsa_full$standards)
  
bsa_full$not_upheld <- coalesce(bsa_full$elect_notupheld, bsa_full$fta_notupheld, bsa_full$paytv_notupheld, bsa_full$radio_notupheld, bsa_full$not_upheld)

bsa_full$upheld <- coalesce(bsa_full$elect_upheld, bsa_full$fta_upheld, bsa_full$paytv_upheld, bsa_full$radio_upheld, bsa_full$upheld)

bsa_full <- select(bsa_full, -c(11:22))
```

Fix factors.WIP.

```{r}
# Recode 'broadcaster' "Newstalk ZB" as the actual Broadcaster

bsa_full$broadcaster <- recode(bsa_full$broadcaster, "Newstalk ZB" = "NZME Radio Ltd")

# Recode 'complaint' numbers as what they have been renumbered to

bsa_full$complaint <- recode(bsa_full$complaint, 
                             "2016-053 (renumbered 048B)" = "2016-048B",
                             "2017-027 (renumbered 018B)" = "2017-018B",
                             "2017-028 (renumbered 018C)" = "2017-018C",
                             "2017-029 (renumbered 018D)" = "2017-018D")

# Recode programme to remove duplicates with minor spelling variations and date variations

bsa_full <- bsa_full |>
  mutate(
    programme = case_when((substr(programme, 1, 6)) %in% c("1 News", "1 NEWS", "One News", "ONE News") ~ "1 News",
                              (substr(programme, 1, 3)) == "Ban" ~ "Ban 1080 Election Programme",
                              (substr(programme, 1, 4)) == "Bhak" ~ "Bhakhde Masley",
                              (substr(programme, 1, 8)) == "Box Seat" ~ "Box Seat",
                              (substr(programme, 1, 9)) %in% c("Breakfast", "TV1 Break") ~ "Breakfast",
                              (substr(programme, 1, 19)) == "Canterbury Mornings" ~ "Canterbury Mornings",
                              (substr(programme, 1, 5)) == "Dasam" ~ "Dasam Granth da Sach",
                              (substr(programme, 1, 13)) == "Early Edition" ~ "Early Edition",
                              (substr(programme, 1, 21)) == "Entertainment Tonight" ~ "Entertainment Tonight",
                              (substr(programme, 1, 7)) == "Fair Go" ~ "Fair Go",
                              (substr(programme, 1, 17)) == "Hauraki Breakfast" ~ "Hauraki Breakfast",
                              (substr(programme, 1, 13)) == "Have You Been" ~ "Have You Been Paying Attention",
                              (substr(programme, 1, 14)) == "Inside the Red" ~ "Inside the Red Arrows",
                              (substr(programme, 1, 5)) == "Magic" ~ "Magic Afternoons",
                              (substr(programme, 1, 14)) == "Morning Report" ~ "Morning Report",
                              (substr(programme, 1, 16)) == "Naked Attraction" ~ "Naked Attraction",
                              (substr(programme, 1, 6)) %in% c("Newshu", "Newhub") ~ "Newshub",
                              (substr(programme, 1, 12)) == "Nine To Noon" ~ "Nine To Noon",
                              (substr(programme, 1, 14)) == "Overnight Talk" ~ "Overnight Talk",
                              (substr(programme, 1, 14)) %in% c("Q+A", "Q +") ~ "Q+A",
                              (substr(programme, 1, 8)) == "RNZ News" ~ "RNZ News",
                              (substr(programme, 1, 11)) == "Seven Sharp" ~ "Seven Sharp",
                              (substr(programme, 1, 16)) == "Shortland Street" ~ "Shortland Street",
                              (substr(programme, 1, 11)) == "The AM Show" ~ "The AM Show",
                              (substr(programme, 1, 12)) == "The Windsors" ~ "The Windsors",
                              .default = programme,
                              )
  )

# Recode 'code' as 'either 'Television' or 'Radio', move to 'tv_radio' column to align with new codebook, then delete 'code' column.

bsa_full$code <- recode(bsa_full$code, 
                             "Free to Air TV" = "Television",
                             "Pay TV" = "Television",
                             "Election Program" = "Television")

bsa_full$tv_radio <- coalesce(bsa_full$code, bsa_full$tv_radio)

bsa_full <- select(bsa_full, -code)

# Remove asterisk and whitespace from columns containing them.

bsa_full <- bsa_full |> 
  mutate(
    standards = gsub("\\*", "", standards),
    not_upheld = gsub("\\*", "", not_upheld),
    upheld = gsub("\\*", "", upheld),
  )

bsa_full$standards <- str_trim(bsa_full$standards, "left")
bsa_full$upheld <- str_trim(bsa_full$upheld, "left")
bsa_full$not_upheld <- str_trim(bsa_full$not_upheld, "left")

# Fix missing values in standards/upheld/not_upheld columns
## Find rows where standards are missing from both 'upheld' and 'not_upheld" columns.

missing <- bsa_full |> 
  filter(is.na(upheld) & is.na(not_upheld))

## Add missing values based on manual BSA decision database lookup

missing[1,10] <- "Discrimination and Denigration"
missing[4, 10] <- "Accuracy" # Complaint 2018-020 
missing[5,10] <- "Balance,Accuracy" # Complaint 2019-006
missing[5,4] <- as.Date("2018-12-17")

## Use case_when to fill in missing values in not_upheld and upheld columns.

missing <- missing |>
  mutate(
    not_upheld = case_when(determination %in% c("Not Upheld", "Declined to Determine 11a", "Declined to Determine 11b") ~ standards,
                           complaint == "2017-101C" ~ "Children's Interests",
                           complaint == "2017-101D" ~ "Children's Interests",
                           complaint == "2019-019B" ~ "Privacy"),
    upheld = case_when(complaint == "2019-019B" ~ "Fairness",
                       complaint == "2017-101C" ~ "Good Taste and Decency",
                       complaint == "2017-101D" ~ "Good Taste and Decency")
  )

## Merge missing dataframe back into bsa_full

bsa_full <- bsa_full |>
  filter(!is.na(upheld) | !is.na(not_upheld)) |> 
  bind_rows(missing)

## Find empty 'standards' values

missing2 <- bsa_full |> 
  filter(is.na(standards))

## Use case_when to fill in missing values in 'standards' columns.

missing2 <- missing2 |> 
  mutate(
    standards = case_when(determination == "Not Upheld" ~ not_upheld,
                          determination == "Upheld" ~ upheld)
  )

## Merge missing standards dataframe back into bsa_full
 
bsa_full <- bsa_full |>
  filter(!is.na(standards)) |> 
  bind_rows(missing2)

```

Create new variable 'complaint_year' based on first 4 characters of 'complaint'.

```{r}
bsa_full$complaint_year <- substr(bsa_full$complaint, 1, 4)
```

Reorder columns.

```{r}
bsa_full <- bsa_full[, c(1,3,5,4,2,14,13,6,7:12)]
```

Swap standards to column names and not_upheld/upheld to values for each complaint.

```{r}
bsa_full <- bsa_full |>
  pivot_longer(cols = c(upheld, not_upheld), values_drop_na = TRUE) |> 
  separate_longer_delim(value, ",") |> 
  pivot_wider(names_from = value, values_from = name)
```

